Covariate analysis of MAYO 50% downsampled reprocessed counts (with CQN normalisation and covariates adjusted)

### 2017-12-22 code/Rmd/MAYO_REPROCESSED_DS.Rmd
Added reprocessing markdown for downsampled Mayo rnaseq data

---
title: "Covariate analysis of MAYO reprocessed counts (with CQN normalisation and covariates adjusted)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')

@@

knit2synapse::knitToFolderEntity(file = "./MAYO_REPROCESSED.Rmd",
                                 parentId = "syn8672415",
                                 entityName = 'MAYO Reprocessed RNASeq (Normalised)')
```

@@

thisFileName <- 'MAYO_REPROCESSED.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD_reprocessing')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
### Data download
#### Obtain count matrix and metadata from synapse.
```{r download.data}
# Get reprocessed sample counts for temporal cortex
COUNT_TC_ID <- 'syn8690799';#'syn7987765';
ALL_USED_IDs <- COUNT_TC_ID
COUNT_TC <- read.table(synGet(COUNT_TC_ID)@filePath, header=T, sep='\t', check.names = F, row.names = 1)

# Get reprocessed sample counts for cerebellum
COUNT_CE_ID <- 'syn8690904';#'syn7513807';
ALL_USED_IDs[length(ALL_USED_IDs)+1] <- COUNT_CE_ID
COUNT_CE <- read.table(synGet(COUNT_CE_ID)@filePath, header=T, sep='\t', check.names = F, row.names = 1)

COUNT = cbind(COUNT_TC, COUNT_CE[rownames(COUNT_TC),])

@@

# Get clinical metadata
METADATA_TC_ID <- 'syn3817650'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_TC_ID
METADATA_TC <- read.table(synGet(METADATA_TC_ID)@filePath,sep='\t',header=T, stringsAsFactors=F) %>%
  tidyr::separate(ID, c('Donor_ID', 'BrainRegion'), sep = '_') %>%
  dplyr::mutate(ID = paste(Donor_ID, BrainRegion, sep = '_'))

# Get clinical metadata
METADATA_CE_ID <- 'syn5223705'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_CE_ID
METADATA_CE <- read.table(synGet(METADATA_CE_ID)@filePath,sep='\t',header=T, stringsAsFactors=F) %>%
  tidyr::separate(SampleID, c('Donor_ID', 'BrainRegion')) %>%
  dplyr::mutate(ID = paste(Donor_ID, BrainRegion, sep = '_')) %>%
  dplyr::rename(Gender = Sex, FLOWCELL = Flowcell)

# Get picard metrics from synapse
METADATA.PICARD_CE_ID <- 'syn8698214';#'syn8380392'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.PICARD_CE_ID
# METADATA.PICARD_CE_FPATHS <- untar(synGet(METADATA.PICARD_CE_ID)@filePath, list = FALSE, exdir = getwd())
# METADATA.PICARD_CE_FPATHS <- untar(synGet(METADATA.PICARD_CE_ID)@filePath, list = TRUE, exdir = getwd())
# METADATA.PICARD_CE_FPATHS = METADATA.PICARD_CE_FPATHS[grep('picard.CombinedMetrics.csv', METADATA.PICARD_CE_FPATHS)]
# METADATA.PICARD_CE <- plyr::ldply(METADATA.PICARD_CE_FPATHS, read.csv)
# 
# METADATA.PICARD_CE$ID = gsub('picard/', '', METADATA.PICARD_CE_FPATHS) %>%
#   gsub('/picard.CombinedMetrics.csv','',.) %>%
#   sapply(function(x){str_split(x, '\\.')[[1]][1]})
METADATA.PICARD_CE <- CovariateAnalysis::downloadFile(METADATA.PICARD_CE_ID) %>%
  dplyr::rename(ID = sample)
colnames(METADATA.PICARD_CE) = gsub('AlignmentSummaryMetrics__','',colnames(METADATA.PICARD_CE))
colnames(METADATA.PICARD_CE) = gsub('RnaSeqMetrics__','',colnames(METADATA.PICARD_CE))

METADATA.PICARD_TC_ID <- 'syn8698211';#'syn8383233'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.PICARD_TC_ID
# METADATA.PICARD_TC_FPATHS <- untar(synGet(METADATA.PICARD_TC_ID)@filePath, list = FALSE, exdir = getwd())
# METADATA.PICARD_TC_FPATHS <- untar(synGet(METADATA.PICARD_TC_ID)@filePath, list = TRUE, exdir = getwd())
# METADATA.PICARD_TC_FPATHS = METADATA.PICARD_TC_FPATHS[grep('picard.CombinedMetrics.csv', METADATA.PICARD_TC_FPATHS)]
# METADATA.PICARD_TC <- plyr::ldply(METADATA.PICARD_TC_FPATHS, read.csv)
# METADATA.PICARD_TC$ID = gsub('picard/', '', METADATA.PICARD_TC_FPATHS) %>%
#   gsub('/picard.CombinedMetrics.csv','',.) %>%
#   sapply(function(x){str_split(x, '\\.')[[1]][1]}) %>%
#   gsub('Aligned','',.)

METADATA.PICARD_TC <- CovariateAnalysis::downloadFile(METADATA.PICARD_TC_ID) %>%
  dplyr::rename(ID = sample)
colnames(METADATA.PICARD_TC) = gsub('AlignmentSummaryMetrics__','',colnames(METADATA.PICARD_TC))
colnames(METADATA.PICARD_TC) = gsub('RnaSeqMetrics__','',colnames(METADATA.PICARD_TC))

@@

tmp = COVARIATES %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Diagnosis'), sep = '\\.') %>%
  group_by(BrainRegion, Diagnosis) %>%
  summarise(count = n()) %>%
  spread(Diagnosis, count)
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

@@

@@

# Set annotations
all.annotations = list(
  dataType = 'mRNA',
  dataSubType = 'geneExp',
  summaryLevel = 'gene',
  assay	 = 'RNAseq',
  assaySubType = 'Reprocessed',

@@

# Code
CODE <- Folder(name = "MAYO Reprocessed RNASeq (Normalised)", parentId = parentId)
annotations(CODE) = all.annotations
CODE <- synStore(CODE)

>>>>

---
title: "Covariate analysis of MAYO 50% downsampled reprocessed counts (with CQN normalisation and covariates adjusted)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')

@@

knit2synapse::knitToFolderEntity(file = "./MAYO_REPROCESSED_DS.Rmd",
                                 parentId = "syn8672415",
                                 entityName = 'MAYO 50percent Down Sampled Reprocessed RNASeq')
```

@@

thisFileName <- 'MAYO_REPROCESSED_DS.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD_reprocessing')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
### Data download
#### Obtain count matrix and metadata from synapse.
```{r download.data}
# Get reprocessed sample counts for cerebellum and temporal cortex
COUNT_ID = 'syn11635112'
ALL_USED_IDs = COUNT_ID
COUNT = read.table(synGet(COUNT_ID)@filePath, header=T, sep='\t', check.names = F, row.names = 1)

@@

# Get clinical metadata
METADATA_TC_ID <- 'syn3817650'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_TC_ID
METADATA_TC <- fread(synGet(METADATA_TC_ID)@filePath,fill=T) %>%
  tidyr::separate(ID, c('Donor_ID', 'BrainRegion'), sep = '_') %>%
  dplyr::mutate(ID = paste(Donor_ID, BrainRegion, sep = '_'))

# Get clinical metadata
METADATA_CE_ID <- 'syn5223705'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_CE_ID
METADATA_CE <- fread(synGet(METADATA_CE_ID)@filePath, fill = T) %>%
  tidyr::separate(SampleID, c('Donor_ID', 'BrainRegion')) %>%
  dplyr::mutate(ID = paste(Donor_ID, BrainRegion, sep = '_')) %>%
  dplyr::rename(Gender = Sex, FLOWCELL = Flowcell)

# Get picard metrics from synapse
METADATA.PICARD_CE_ID <- 'syn8698214';#'syn8380392'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.PICARD_CE_ID
METADATA.PICARD_CE <- CovariateAnalysis::downloadFile(METADATA.PICARD_CE_ID) %>%
  dplyr::rename(ID = sample)
colnames(METADATA.PICARD_CE) = gsub('AlignmentSummaryMetrics__','',colnames(METADATA.PICARD_CE))
colnames(METADATA.PICARD_CE) = gsub('RnaSeqMetrics__','',colnames(METADATA.PICARD_CE))

METADATA.PICARD_TC_ID <- 'syn8698211';#'syn8383233'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.PICARD_TC_ID
METADATA.PICARD_TC <- CovariateAnalysis::downloadFile(METADATA.PICARD_TC_ID) %>%
  dplyr::rename(ID = sample)
colnames(METADATA.PICARD_TC) = gsub('AlignmentSummaryMetrics__','',colnames(METADATA.PICARD_TC))
colnames(METADATA.PICARD_TC) = gsub('RnaSeqMetrics__','',colnames(METADATA.PICARD_TC))

@@

tmp = COVARIATES %>%
  tidyr::separate(BrainRegion.Diagnosis, c('BrainRegion', 'Diagnosis'), sep = '\\.') %>%
  dplyr::group_by(BrainRegion, Diagnosis) %>%
  dplyr::count() %>%
  tidyr::spread(Diagnosis, n)
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

@@

# Fit linear model using new weights dplyr::count()and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS$E,
            weights = VOOM.WEIGHTS$weights,
            design = VOOM.WEIGHTS$design,
            block = COVARIATES$Donor_ID, 
            correlation = correlation$cor)

@@

# Set annotations
all.annotations = list(
  dataType = 'mRNA',
  dataSubType = 'geneExp',
  summaryLevel = 'gene',
  assay	 = 'RNAseq',
  assaySubType = 'Reprocessed down sampled',

@@

# Code
CODE <- Folder(name = "MAYO 50percent Down Sampled Reprocessed RNASeq", parentId = parentId)
annotations(CODE) = all.annotations
CODE <- synStore(CODE)
